Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-02-05 paper/paper.tex
added data for other_arrow_plot.py and also mannik section edits to paper Tested everything on node32.

In general process results in extremely robust oscillatory behavoir of
polar selection and oscillation in a wide variety of cell shapes.

@@

\section{Theory}
Stability analysis of the five differential equations in an infinite
slab show a characteristic half wavelength in perturbative spatial
oscillations at which the system becomes unstable.  This length is
dependent upon slab thickness as shown in
Figure~\ref{fig:stability-analysis}.  When increasing the lengths and
widths of our simulated flattened cells, the steady, two pole
oscillations seem to break down when the the shortest distance across
the cell is approximately equal to the characteristic half wavelength
at the particular thickness at which we simulate (our simulations of cells
with thickness $0.25\mu m$ corresponds to half wavelengths of $1.91\mu
m$).  Huang~\cite{huang2003dynamic} performs a similar analysis for a
cylindrical shape and finds characteristic half wavelengths of $2\mu
m$.

\subsection{Randst shape}
\fixme{Cheat sheet for the shapes, need to put these into a big figure
  from the sections output: 96=star shape, 97=sideways sideways
  flying saucer pancake, 98=manniks squished cell, 99=an A}

\subsubsection{Randst-mannik shape}
\begin{figure}
  \includegraphics[width=2.8cm]{../mannik-2.png}
  \includegraphics[width=2.8cm]{../mannik-1.png}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1300-2000-9400-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1100-1100-9500-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1800-2500-9400-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1800-1800-9500-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-2300-3300-9400-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-2800-2800-9500-1500.pdf}
  \caption{Arrow plots for our simulation and mannik's experimental}
  \label{arrow-compare-mannik}
\end{figure}

Figure~\ref{arrow-compare-mannik} shows a comparison of time maxima
plots with similar plots published by mannik et al.  The sequence of
arrows show a series of global maxima in space and local maxima in
time at their heads, numbered by the sequence in which they occur.
They function as a trace of the path that the traveling maxima take.
We have created for each of the two shapes published by Mannik a
closely resembling shape (shown directly below Mannik's) that is in
all dimensions the same size as that reported in their experiment.  We
have also simulated longer and wider versions of these same shapes
(keeping the cell thickness the same).  Shown in
Figure~\ref{arrow-compare-mannik} are versions that have 1.3 and 2.0
times the length and width dimensions.  These plots show
that there are certain sizes at which the steady, two pole
oscillations seem to break down, and indeed these lengths seem to be
roughly those estimated by our 2D stability analysis.  \fixme{Waiting
  to rewrite this section when more data, at more sizes, becomes
  available} The plots of the left shape show that much of the
preffered locations for the proteins to maximize are similar between
this experimental and simulated shape.  The maxima in the top center
of the cell does not seem to appear in the experimental version.  Also
in both of the simulations on the left there don't seem to be maxima
at membrane regions with negative radii of curvature while the
experimental shows a maxima in the bottom center region right at one
of these regions.  The shape on the right compares less successfully.
Here the simulation shows a limited number of maxima locations that
appear to be quite regular, where mannik's experimental results show a
much more chaotic system.  In particular, the maxima location on the
right wall seems to be very robust.  It acts as one of the poles do in
the pill shapes.  \fixme{I'm now running a sim with a slightly
  different shape to see result} The simulated version has found a
robust two pole pattern, similar to that seen in other shapes.

\subsection{Triangular Shapes}
\begin{figure}
  %%\includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-400-400-400-1500}
  %%\includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-500-300-500-1500}
  %%\includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-600-480-360-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of an equilateral (top), iscoscolese (middle), and 3-4-5
    (bottom) triangle.  The vertical axis shows stacked the total
    number of proteins that are of four different compound stages and
    in the different sections of the cell.}
  \label{box-triangle}
\end{figure}

>>>>

@@

\section{Stability Analysis and Mannik's Experimental Shapes}
In general we have found that the protein reaction sequence discussed
above results in extremely robust oscillatory behavoir of polar
selection and oscillation in a wide variety of cell shapes.  However
we have found that this robustness is heavily dependent upon cell
size.  Two pole, regular oscillatory behavoir is found in cells which
are long enough in one dimension for the reaction chain to be unstable
but short enough in the perpendicular dimension to be stable.  The
limits are seen when both dimensions are too short to be unstable, so
that initial assymetry in protein concentration relaxes into a
motionless steady state, or when both are too great, so that the
proteins exhibit much more complex and varied behavoir.

In order to explain this theoretically, we have performed stability
analysis of the five differential equations in an infinite slab that
shows a characteristic half wavelength in perturbative spatial
oscillations at which the system becomes unstable.  This length is
dependent upon slab thickness as shown in
Figure~\ref{fig:stability-analysis}.  When increasing the lengths and
widths of our simulated flattened cells, the steady, two pole
oscillations seem to break down when the the shortest distance across
the cell is approximately equal to the characteristic half wavelength
at the particular thickness at which we simulate (our simulations of
cells with thickness $0.25\mu m$ corresponds to half wavelengths of
\fixme{check next number} $1.91\mu m$).  Huang~\cite{huang2003dynamic}
performs a similar analysis for a cylindrical shape and finds
characteristic half wavelengths of $2\mu m$.

\begin{figure}
  \includegraphics[width=2.8cm]{../mannik-2.png}
  \includegraphics[width=2.8cm]{../mannik-1.png}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1300-2000-9400-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1100-1100-9500-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1800-2500-9400-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1800-1800-9500-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-2300-3300-9400-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-2800-2800-9500-1500.pdf}
  \caption{Arrow plots for our simulation and mannik's experimental}
  \label{arrow-compare-mannik}
\end{figure}

Figure~\ref{arrow-compare-mannik} shows plots of total MinD
concentration maxima that are global in space and local in time.  The
tip of an arrow represents one of these maxima and the adjacent arrow
(whose tail is touching the previous arrow's tip) points to the next
such maxima in time.  Mannik \emph{et al.} have published similar
plots of two of their cells, and we have attempted to replicate the
shape of these cells in our simulation in order to compare.  We find
that when we replicate the size as well as shape of their cells as
near as possible, we do not see the disordered protein behavoir that
they observe.  Instead we see a robust ability of the cell to find two
prefered pole locations and to oscillate maxima back and forth very
steadily between them.  However, when we keep the same shape but
increase the size, we see that there is a critical size at which the
protein dynamics lose their order and become very complex (showing
hints of prefered wall maxima locations), similar to what Mannik sees
experimentally.  We find that this critical size matches up well with
our stability analysis, in that the critical sizes correspond to sizes
for which the shorter of the two distances across the cell is roughly
equal to the analysis' half wavelength found for this cell thickness.

\fixme{I'm also waiting for data to come in that will hopefully show
  cells so small that no oscillations take place.  We can put this in
  as well}

\section{Triangular Shapes}
\begin{figure}
  %%\includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-400-400-400-1500}
  %%\includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-500-300-500-1500}
  %%\includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-600-480-360-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of an equilateral (top), iscoscolese (middle), and 3-4-5
    (bottom) triangle.  The vertical axis shows stacked the total
    number of proteins that are of four different compound stages and
    in the different sections of the cell.}
  \label{box-triangle}
\end{figure}
